---
title: "Collaboration KPI Overview"
format: html
editor: visual
---
```{r initialize, echo=FALSE, message=FALSE, warning=FALSE}
library(fontawesome)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(ggraph)
library(stringr)
library(tidygraph)
library(httr2)
library(purrr)
library(digest)
library(stringdist)
library(cli)

source("./Functions_Collaboration_ABPA.R")

theIteration <- 8
numPubsToProcess <- 3000
keepYears <- c(2025)
keepStatus <- c("Completed/Published")
orcidEval <- FALSE

#tempPubs <- readRDS(paste0("../R//someDataMonday", theIteration, ".rds"))
tempPubs <- readRDS("../R/temp1.rds")          
#assignt(paste0("someDataMonday", theIteration, ".rds"), tempPubs)
allPubs <- read.csv("../Data/testPubs2024-2025.csv")
deptList <- readRDS("../Data/deptList.rds")
usePubs_verified_doi <- readRDS("../usePubs_verified_doi.rds")
additionalDOI <- readRDS("../Data/additionalDOI.rds")
testPubsAll <- readRDS("../Data/testPubsAll.rds")
testPubsAll <- testPubsAll |>
  filter(Year...22 %in% keepYears)


```

## Collaboration KPI Overview

This document provides an overview of the development of collaboration graphics and provides a status update for the project.

Objectives:

-  Provide information about number of publications within a year range.

-  Identify the number of papers that are collaborative (among any institutions).

-  For all publications, determine the

author list

author institution

author department

-  Filter down to only TAMU authors

-  Determine collaboration type at level of College

-  Determine collaboration type at level of Department

-  Circle packing figures

-  Extend to grants awarded and grants applied

Current Status:


Process

`r fontawesome::fa("check", fill = "green")` Retrieve 2025 Publication Data from Interfolio F180.
```{r pubNumbers1, echo=FALSE, message=FALSE, warning=FALSE}
#| output: asis

library(dplyr)
library(glue)
library(htmltools)
library(purrr)

# 1. Calculate status counts while filtering out NA values
status_counts <- testPubsAll %>%
  filter(!is.na(Status)) %>%
  count(Status)

# 2. Generate individual HTML cards dynamically
cards_html <- map2_chr(status_counts$Status, status_counts$n, function(status_name, count) {
  glue('
  <div style="
    flex: 1;
    min-width: 180px;
    background-color: #f8f9fa;
    border: 2px solid #e9ecef;
    border-left: 6px solid #500000;
    border-radius: 8px;
    padding: 20px 24px;
    box-shadow: 0 2px 4px rgba(0,0,0,0.05);
  ">
    <div style="
      font-size: 0.9rem;
      font-weight: 600;
      color: #495057;
      margin-bottom: 8px;
      text-transform: uppercase;
      letter-spacing: 0.5px;
    ">
      {status_name}
    </div>
    <div style="
      font-size: 2.5rem;
      font-weight: 800;
      color: #500000;
      line-height: 1;
    ">
      {format(count, big.mark = ",")}
    </div>
  </div>
  ')
})

# 3. Wrap all generated cards in a flex container
wrapper_html <- glue('
<div style="display: flex; gap: 16px; flex-wrap: wrap; margin: 15px 0; font-family: sans-serif;">
  {paste(cards_html, collapse = "\n")}
</div>
')

# 4. Print HTML output for Quarto / R Markdown
cat("\n\n")
cat(as.character(HTML(wrapper_html)))

```


`r fontawesome::fa("check", fill = "green")` Filter to only consider the completed/published papers.
```{r pubNumbers3, echo=FALSE, message=FALSE, warning=FALSE}
#| output: asis

testPubsFiltered <- testPubsAll %>%
  filter(Status %in% keepStatus)

status_counts <- testPubsFiltered %>%
  filter(!is.na(Status)) %>%
  count(Status)

# 2. Generate individual HTML cards dynamically
cards_html <- map2_chr(status_counts$Status, status_counts$n, function(status_name, count) {
  glue('
  <div style="
    flex: 1;
    min-width: 180px;
    background-color: #f8f9fa;
    border: 2px solid #e9ecef;
    border-left: 6px solid #500000;
    border-radius: 8px;
    padding: 20px 24px;
    box-shadow: 0 2px 4px rgba(0,0,0,0.05);
  ">
    <div style="
      font-size: 0.9rem;
      font-weight: 600;
      color: #495057;
      margin-bottom: 8px;
      text-transform: uppercase;
      letter-spacing: 0.5px;
    ">
      {status_name}
    </div>
    <div style="
      font-size: 2.5rem;
      font-weight: 800;
      color: #500000;
      line-height: 1;
    ">
      {format(count, big.mark = ",")}
    </div>
  </div>
  ')
})

# 3. Wrap all generated cards in a flex container
wrapper_html <- glue('
<div style="display: flex; gap: 16px; flex-wrap: wrap; margin: 15px 0; font-family: sans-serif;">
  {paste(cards_html, collapse = "\n")}
</div>
')

# 4. Print HTML output for Quarto / R Markdown
cat("\n\n")
cat(as.character(HTML(wrapper_html)))

```

How many pubs have no DOI information?

```{r pubNumbers4, echo=FALSE, message=FALSE, warning=FALSE}
#| output: asis

testPubsFilteredNoDOI <- testPubsFiltered |>
  filter(is.na(DOI)) |>
  distinct()


status_counts <- testPubsFilteredNoDOI %>%
  filter(!is.na(Status)) %>%
  count(Status)

# 2. Generate individual HTML cards dynamically
cards_html <- map2_chr(status_counts$Status, status_counts$n, function(status_name, count) {
  glue('
  <div style="
    flex: 1;
    min-width: 180px;
    background-color: #f8f9fa;
    border: 2px solid #e9ecef;
    border-left: 6px solid #500000;
    border-radius: 8px;
    padding: 20px 24px;
    box-shadow: 0 2px 4px rgba(0,0,0,0.05);
  ">
    <div style="
      font-size: 0.9rem;
      font-weight: 600;
      color: #495057;
      margin-bottom: 8px;
      text-transform: uppercase;
      letter-spacing: 0.5px;
    ">
      {status_name}
    </div>
    <div style="
      font-size: 2.5rem;
      font-weight: 800;
      color: #500000;
      line-height: 1;
    ">
      {format(count, big.mark = ",")}
    </div>
  </div>
  ')
})

# 3. Wrap all generated cards in a flex container
wrapper_html <- glue('
<div style="display: flex; gap: 16px; flex-wrap: wrap; margin: 15px 0; font-family: sans-serif;">
  {paste(cards_html, collapse = "\n")}
</div>
')

# 4. Print HTML output for Quarto / R Markdown
cat("\n\n")
cat(as.character(HTML(wrapper_html)))

```

`r fontawesome::fa("check", fill = "green")` Retrieve DOI using author, title and journal title from CrossRef & OpenAlex.
```{r retrieveDOI, echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE}

additionalDOI <- batch_get_dois(testPubsFilteredNoDOI, max_dist = 0.40)

```

```{r pubNumbers6, echo=FALSE, message=FALSE, warning=FALSE}
#| output: asis

# 1. Calculate status counts while filtering out NA values
status_counts <- additionalDOI %>%
  count(doi_source)

# 2. Generate individual HTML cards dynamically
cards_html <- map2_chr(status_counts$doi_source, status_counts$n, function(status_name, count) {
  glue('
  <div style="
    flex: 1;
    min-width: 180px;
    background-color: #f8f9fa;
    border: 2px solid #e9ecef;
    border-left: 6px solid #500000;
    border-radius: 8px;
    padding: 20px 24px;
    box-shadow: 0 2px 4px rgba(0,0,0,0.05);
  ">
    <div style="
      font-size: 0.9rem;
      font-weight: 600;
      color: #495057;
      margin-bottom: 8px;
      text-transform: uppercase;
      letter-spacing: 0.5px;
    ">
      {status_name}
    </div>
    <div style="
      font-size: 2.5rem;
      font-weight: 800;
      color: #500000;
      line-height: 1;
    ">
      {format(count, big.mark = ",")}
    </div>
  </div>
  ')
})

# 3. Wrap all generated cards in a flex container
wrapper_html <- glue('
<div style="display: flex; gap: 16px; flex-wrap: wrap; margin: 15px 0; font-family: sans-serif;">
  {paste(cards_html, collapse = "\n")}
</div>
')

# 4. Print HTML output for Quarto / R Markdown
cat("\n\n")
cat(as.character(HTML(wrapper_html)))
```

`r fontawesome::fa("check", fill = "green")` Join found DOI data back into original data frame.

```{r joinFound, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
skinnyDOI <- additionalDOI |>
  select(all_of(c("UID", "Last Name", "found_doi", "doi_source")))

testPubsFiltered2 <- testPubsFiltered |>
  left_join(skinnyDOI, by=c("UID", "Last Name")) |>
  mutate(DOI=case_when(!is.na(found_doi) ~ found_doi,
                       TRUE ~ DOI))

```

Total Number of Pubs with DOI
```{r pubNumbers7, echo=FALSE, message=FALSE, warning=FALSE}
#| output: asis

testPubsFilteredNoDOI <- testPubsFiltered2 |>
  filter(!is.na(DOI)) |>
  distinct()


status_counts <- testPubsFilteredNoDOI %>%
  filter(!is.na(Status)) %>%
  count(Status)

# 2. Generate individual HTML cards dynamically
cards_html <- map2_chr(status_counts$Status, status_counts$n, function(status_name, count) {
  glue('
  <div style="
    flex: 1;
    min-width: 180px;
    background-color: #f8f9fa;
    border: 2px solid #e9ecef;
    border-left: 6px solid #500000;
    border-radius: 8px;
    padding: 20px 24px;
    box-shadow: 0 2px 4px rgba(0,0,0,0.05);
  ">
    <div style="
      font-size: 0.9rem;
      font-weight: 600;
      color: #495057;
      margin-bottom: 8px;
      text-transform: uppercase;
      letter-spacing: 0.5px;
    ">
      {status_name}
    </div>
    <div style="
      font-size: 2.5rem;
      font-weight: 800;
      color: #500000;
      line-height: 1;
    ">
      {format(count, big.mark = ",")}
    </div>
  </div>
  ')
})

# 3. Wrap all generated cards in a flex container
wrapper_html <- glue('
<div style="display: flex; gap: 16px; flex-wrap: wrap; margin: 15px 0; font-family: sans-serif;">
  {paste(cards_html, collapse = "\n")}
</div>
')

# 4. Print HTML output for Quarto / R Markdown
cat("\n\n")
cat(as.character(HTML(wrapper_html)))

```

`r fontawesome::fa("check", fill = "green")` Verify DOIs with CrossRef.

```{r verifyDOI, echo=FALSE, message=FALSE, warning=FALSE, eval=FALSE}
testPubsFiltered2 <- testPubsFiltered2 |>
  mutate(doi=DOI,
         uid=UID,
         title=Title,
         author=`Last Name`)
usePubs_verified_doi <- verify_doi_batch(testPubsFiltered2)
```

```{r card6, echo=FALSE, message=FALSE, warning=FALSE}
#| output: asis

# 1. Calculate status counts while filtering out NA values
status_counts <- usePubs_verified_doi %>%
  count(status)

wrapper_html <- make_status_card_html(status_counts, displayCountsForField="status")
#print(status_counts)
# 4. Print HTML output for Quarto / R Markdown
knitr::asis_output(wrapper_html)
```


`r fontawesome::fa("check", fill = "green")` Parse author field and separate into a single row for each author using openalex.

```{r openalexR, echo=FALSE, eval=FALSE, message=FALSE, warning=FALSE}

temp1 <- expand_authors_by_openalex(usePubs_verified_doi)
saveRDS(temp1, "temp1.rds")
```


```{r crossref, echo=FALSE, eval=FALSE, message=FALSE, warning=FALSE}
# currently not used.  openalex produces cleaner, better results.
temp1 <- collaboration_abpa_kpi(numPubs=numPubsToProcess,
                                previouslyDoneBatch=tempPubs,
                                )
saveRDS(temp1, "temp1.rds")
```



For authors with missing affiliation, use check orcid id to determine affiliation.

```{r orcid, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
data_after_orcid <- temp1
if(orcidEval){
data_after_orcid <- check_orcid_for_affiliation(temp1, doEval=orcidEval)
}
saveRDS(data_after_orcid, "data_after_orcid.rds")
```

Parse affiliation to find TAMU authors
```{r tamuAuthors, echo=FALSE, eval=FALSE, message=FALSE, warning=FALSE}
data_after_orcid <- data_after_orcid %>%
  mutate(affiliation=author_affiliation)
data_after_tamu <- addField_TAMU(data_after_orcid)
saveRDS(data_after_tamu, "data_after_tamu.rds")
#data_after_tamu <- addField_TAMU(data_after_orcid)
```

Parse affiliation to determine Department and College
```{r deptAffiliation, echo=FALSE, eval=FALSE, message=FALSE, warning=FALSE}

data_after_dept <- cbind(data_after_tamu, match_phrase_fuzzy(data_after_tamu, deptList=deptList))
saveRDS(data_after_dept, "data_after_dept.RDS")
```



Filter out non-tamu authors
```{r tamuAuthors1, echo=FALSE, eval=FALSE, message=FALSE, warning=FALSE}
data_tamu_authors <- data_after_dept |>
  filter(!is.na(TAMU)) |>
  filter(TAMU==TRUE)

```

Identify collaborations among tamu authors
```{r tamuCollaborations, echo=FALSE, eval=FALSE, message=FALSE, warning=FALSE}
data_tamu_authors <- data_tamu_authors %>%
  mutate(uid=UID)
data_tamu_collabs <- determine_tamu_collaborations(data_tamu_authors)
saveRDS(data_tamu_collabs, "data_tamu_collabs.rds")

```

Using uid as the key, join results back to original dataset

```{r pubNumbers, echo=FALSE, message=FALSE, warning=FALSE}
#| output: asis

library(glue)
library(htmltools)
data_tamu_collabs <- readRDS("data_tamu_collabs.rds")

pubStats <- filterPubsFromF180(allPubs, data_tamu_collabs)

card_html <- glue('
<div style="display: flex; gap: 16px; flex-wrap: wrap; margin: 15px 0; font-family: sans-serif;">
<!-- Total Publications Card -->
<div style="
flex: 1;
min-width: 180px;
background-color: #f8f9fa;
border: 2px solid #e9ecef;
border-left: 6px solid #500000;
border-radius: 8px;
padding: 20px 24px;
box-shadow: 0 2px 4px rgba(0,0,0,0.05);
">
<div style="
font-size: 0.9rem;
font-weight: 600;
color: #495057;
margin-bottom: 8px;
text-transform: uppercase;
letter-spacing: 0.5px;
">
Total Publications
</div>
<div style="
font-size: 2.5rem;
font-weight: 800;
color: #500000;
line-height: 1;
">
{format(pubStats$totalNumPubs, big.mark = ",")}
</div>
</div>

<!-- Processed Publications Card -->
<div style="
flex: 1;
min-width: 180px;
background-color: #f8f9fa;
border: 2px solid #e9ecef;
border-left: 6px solid #0056b3;
border-radius: 8px;
padding: 20px 24px;
box-shadow: 0 2px 4px rgba(0,0,0,0.05);
">
<div style="
font-size: 0.9rem;
font-weight: 600;
color: #495057;
margin-bottom: 8px;
text-transform: uppercase;
letter-spacing: 0.5px;
">
Processed
</div>
<div style="
font-size: 2.5rem;
font-weight: 800;
color: #0056b3;
line-height: 1;
">
{format(pubStats$processedPubs, big.mark = ",")}
</div>
</div>

<!-- Single Author Publications Card -->
<div style="
flex: 1;
min-width: 180px;
background-color: #f8f9fa;
border: 2px solid #e9ecef;
border-left: 6px solid #2B5235;
border-radius: 8px;
padding: 20px 24px;
box-shadow: 0 2px 4px rgba(0,0,0,0.05);
">
<div style="
font-size: 0.9rem;
font-weight: 600;
color: #495057;
margin-bottom: 8px;
text-transform: uppercase;
letter-spacing: 0.5px;
">
Single Author
</div>
<div style="
font-size: 2.5rem;
font-weight: 800;
color: #2B5235;
line-height: 1;
">
{format(pubStats$singleAuthorPubs, big.mark = ",")}
</div>
</div>
</div>
')

# Print with explicit leading whitespace and raw HTML class specification
cat("\n\n")
cat(as.character(HTML(card_html)))

```

Create circle packing figure for all publications to distinguish single author from multiple author papers (one circle for each).

```{r circlePacking, echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}

# edges are parent-child relationships (hierarchy) in the format of a data frame with
# from and to collumns.  No numbers, only the hierarchy,
# e.g. from TAMU to AG
# e.g. from AG to AG:AT

# nodes are values for the leaves (individual circles that are sized proportionally)
# a data frame with collumns "name" and "value".  Terminal leaf values of NA or 0 are
# not allowed.

make_edges_college <- function(inData){
  # make the edges dataframe consisting of to from and to columns
  # from collumns should be TAMU and each Department
  # to collumns should be Departments and collaboration types e.g. (GEOG:ATMO)
  #
  unique_colleges <- unique(unlist(stringr::str_split(inData$college_collab, ":")))
  unique_colleges <- unique_colleges[!is.na(unique_colleges)]
  unique_departments <- unique(unlist(stringr::str_split(inData$collab, ":")))
  unique_dept_collabs <- unique(inData$collab)
  
  edges <- data.frame(
    from = c(
      rep("TAMU", length(unique_colleges)),
      rep(unique_colleges, length(unique(inData$college_collab)))
    ),
    to = c(
      unique_colleges,
      rep(unique(inData$college_collab), length(unique_colleges))
    )
  )
  edges
}  #Not currently used
make_nodes_college <- function(inData){
  unique_colleges <- unique(unlist(stringr::str_split(inData$college_collab, ":")))
  unique_colleges <- unique_colleges[!is.na(unique_colleges)]
  unique_departments <- unique(unlist(stringr::str_split(inData$collab, ":")))
  unique_dept_collabs <- unique(inData$collab)
  nodes <- data.frame(
    name=c("TAMU", unique_colleges),
    value=c(0,
            rep(0, length(unique_colleges))
    
    )
  )
  # create a table with number of times a college has a collab pub.
  # count the numer of individual ATs for instance in the inData$
  nodes
}  #Not currently used

link_colleges_departments <- function(inData) {
  outData <- inData |>
    select("COLLEGE_CODE", "Department") |>
    filter(!is.na(COLLEGE_CODE)) |>
    arrange(COLLEGE_CODE, Department) |>
    distinct()
  
  names(outData) <- c("from", "to")
  outData
}  #Not currently used

# 1. Edges Data Frame
#############################################
# Edges is the hierarchical structure.      #
# Contains a from and a to column           #
# Every hierarhica; combination must be     #
# present.                                  #
# TAMU to each college.                     #
# Each College to consituent departments.   #
#############################################

create_edges_df <- function(data, root_name = "TAMU", min_size = 1) {
  # 1. Get filtered collaboration pairs meeting the minimum frequency threshold
  pairs <- data %>%
    filter(!is.na(college_collab) & college_collab != "") %>%
    count(college_collab, name = "size") %>%
    filter(size >= min_size)
  
  # Handle empty result if threshold is too high
  if (nrow(pairs) == 0) {
    return(tibble(from = character(), to = character()))
  }
  
  # 2. Level 2 Edges: College -> College Collab Pairs (e.g., AG -> AG:AG, EN -> AG:EN)
  edges_l2 <- pairs %>%
    separate(college_collab, into = c("c1", "c2"), sep = ":", remove = FALSE) %>%
    pivot_longer(cols = c(c1, c2), values_to = "from") %>%
    select(from, to = college_collab) %>%
    distinct()
  
  # 3. Level 1 Edges: Root -> Active Colleges
  edges_l1 <- tibble(
    from = root_name,
    to = unique(edges_l2$from)
  )
  
  # Combine hierarchy levels
  bind_rows(edges_l1, edges_l2)
}

create_nodes_df <- function(data, edges_df, root_name = "TAMU") {
  # 1. Get count for leaf nodes present in the filtered edges
  leaf_counts <- data %>%
    filter(!is.na(college_collab) & college_collab != "") %>%
    count(college_collab, name = "size") %>%
    filter(college_collab %in% edges_df$to)
  
  # 2. Get all unique nodes dynamically present in the edges graph
  all_nodes <- unique(c(edges_df$from, edges_df$to))
  
  # 3. Build nodes data frame with hierarchy groups and sizes
  nodes_df <- tibble(name = all_nodes) %>%
    left_join(leaf_counts, by = c("name" = "college_collab")) %>%
    mutate(
      # Parent nodes (TAMU & Colleges) have size = 0 so ggraph aggregates child sizes
      size = ifelse(is.na(size), 0, size),
      
      # Assign structural levels/groups for styling/coloring
      group = case_when(
        name == root_name ~ "Root",
        !str_detect(name, ":") ~ "College",
        TRUE ~ "Collaboration"
      )
    )
  
  return(nodes_df)
}

create_edges_df_orig <- function(data, root_name = "TAMU") {
  # Get all distinct college_collab values
  pairs <- data %>%
    filter(!is.na(college_collab) & college_collab != "") %>%
    distinct(college_collab)
  
  # Level 2 Edges: College -> College Collab Pairs (e.g., AG -> AG:AG, EN -> AG:EN)
  edges_l2 <- pairs %>%
    separate(college_collab, into = c("c1", "c2"), sep = ":", remove = FALSE) %>%
    pivot_longer(cols = c(c1, c2), values_to = "from") %>%
    select(from, to = college_collab) %>%
    distinct()
  
  # Level 1 Edges: Root -> Individual Colleges (e.g., TAMU -> AG, TAMU -> EN)
  edges_l1 <- tibble(
    from = root_name,
    to = unique(edges_l2$from)
  )
  
  # Combine both hierarchy levels
  bind_rows(edges_l1, edges_l2)
}

edges.auto <- create_edges_df(data_tamu_collabs, min_size =4)

create_nodes_df_orig <- function(data, edges_df, root_name = "TAMU") {
  # 1. Count collaboration frequencies for leaf node sizes
  leaf_counts <- data %>%
    filter(!is.na(college_collab) & college_collab != "") %>%
    count(college_collab, name = "size")
  
  # 2. Get all unique nodes from the edges data frame
  all_nodes <- unique(c(edges_df$from, edges_df$to))
  
  # 3. Build nodes data frame with hierarchy groups and sizes
  nodes_df <- tibble(name = all_nodes) %>%
    left_join(leaf_counts, by = c("name" = "college_collab")) %>%
    mutate(
      # Parent nodes (TAMU & Colleges) have size = 0 so ggraph aggregates child sizes
      size = ifelse(is.na(size), 0, size),
      
      # Assign structural levels/groups for styling/coloring
      group = case_when(
        name == root_name ~ "Root",
        !str_detect(name, ":") ~ "College",
        TRUE ~ "Collaboration"
      )
    )
  
  return(nodes_df)
}

nodes.auto <- create_nodes_df(data_tamu_collabs, edges.auto)

# 3. Create Graph
graph <- tbl_graph(nodes = nodes.auto, edges = edges.auto)

# 4. Plot with leaf filtering
ggraph(graph, layout = "circlepack", weight = size) +
  # Layer 1: Background containers (Root & Colleges)
  geom_node_circle(
    aes(filter = !leaf), 
    fill = "cyan", 
    color = "grey70", 
    linewidth = 0.5
  ) +
  # Layer 2: Foreground leaf circles (Collaboration pairings)
  geom_node_circle(
    aes(fill = group, filter = leaf), 
    color = "white", 
    linewidth = 0.5
  ) +
  # Layer 3: Labels on leaf nodes with collaborations
  geom_node_text(
    aes(label = name, filter = leaf), 
    size = 3, 
    fontface = "bold"
  ) +
  theme_void() +
  coord_fixed()

```



